---
title: "OmicsEV: A tool for large scale omics datasets evaluation"
author: "Bo Wen"
date: "`r Sys.Date()`"
bibliography: OmicsEV.bib
output: 
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{00 OmicsEV introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r style, echo=FALSE, results="asis", message=FALSE}
knitr::opts_chunk$set(tidy = FALSE,message = FALSE)
```



```{r echo=FALSE, results="hide"}
library("BiocStyle")
BiocStyle::markdown()
```




```{r echo=FALSE,warning=FALSE}
suppressPackageStartupMessages(library("OmicsEV"))
suppressPackageStartupMessages(library("R.utils"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("kableExtra"))
suppressPackageStartupMessages(library("formattable"))
```


# Introduction

High-throughput technologies such as RNA-Seq and mass spectrometry-based 
proteomics are increasingly being applied to large sample cohorts, which 
creates vast amount of quantitative data for genes and proteins. Many algorithms, 
software, and pipelines have been developed to analyze these data. However, 
how to select optimal algorithms, software, and parameters for analyzing a 
specific omics dataset remains a significant challenge. To address this 
challenge, we have developed an R package named `OmicsEV`, which is dedicated to 
compare and evaluate different data matrices generated from the same omics 
dataset using different tools, algorithms, or parameter settings. In `OmicsEV`,
we have implemented more than 15 evaluation metrics and all the evaluation 
results are included in an HTML-report for intuitive browsing. OmicsEV is easy 
to install and use. Only one function is needed to perform the whole evaluation 
process. A GUI based on R shiny is also implemented.

# Example data

An example data can be downloaded at https://github.com/bzhanglab/OmicsEV. This
example contains 6 data matrices generated from the same proteomics dataset 
using different normalization methods. In addition, an RNA-Seq data matrix and a
sample list are also included.

# Running OmicsEV

### Preparing inputs

### Running evaluation process

In `OmicsEV`, Only one function (**run_omics_evaluation**) is needed to 
perform the whole evaluation process. An example is showing below:

```{r eval=FALSE, echo=TRUE}
library(OmicsEV)
run_omics_evaluation(data_dir = "datasets/",
                     sample_list = "sample_list.tsv",
                     x2 = "protein.tsv",
                     cpu=6,
                     data_type="gene",
                     class_for_ml="sample_ml.tsv")
```

In general, only a few parameters have to be set:

* **data_dir**: a folder contains the datasets in tsv format to be evaluated. All 
datasets must be the same format. In these files, the first column must be gene
ID or protein ID. The expression value of gene or protein must be in non-log 
scale. An example input dataset looks like below:

```{r eval=TRUE, echo=FALSE, results='asis'}
example_data <- read.delim(system.file("extdata/example_input_datasets.tsv",
                                       package = "OmicsEV"),
                           stringsAsFactors = FALSE)
kable(example_data,digits = 3,caption="An example of input dataset") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

* **sample_list**: a file in tsv format contains sample class, batch and order 
information. This file must contain sample ID, sample class, batch and order 
information. An example sample list file looks like below:

```{r eval=TRUE, echo=FALSE, results='asis'}
example_data <- read.delim(system.file("extdata/example_sample_list.tsv",
                                       package = "OmicsEV"),
                           stringsAsFactors = FALSE)
kable(example_data,digits = 3,caption="An example of sample list") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

* **data_type**: the quantification data type in folder data_dir: protein, 
gene. Default is protein.

All other parameters are optional. When input datasets for parameter 
**data_dir** are protein expression data and users also have gene expression 
data for the same samples, users can set parameter **x2** as a file contains 
the gene expression data in tsv format, and vice versa. If parameter **x2** is 
not NULL, sample wise and gene wise correlation analysis will be performed. See `?run_omics_evaluation` for a more in-depth description of all its arguments. 

When the function is finished successfully, an HTML-based report that contains 
different evaluation metrics will be generated. Example reports are available 
at https://github.com/bzhanglab/OmicsEV.


# Evaluation metrics implemented in OmicsEV

So far, more than 15 evaluation metrics have been implemented in `OmicsEV` and 
the evaluation result is organized in the following structure:

1. Introduction
2. Overview
3. Data depth
   a. Study-wise (#identified features, #quantifiable features)
   b. Sample-wise
   c. Missing value distribution (Non-missing value percentage in the data table)
4. Data normalization
   a. Boxplot (Data distribution similarity)
   b. Density plot
5. Batch effect
   a. Silhouette width (silhouette width)
   b. PCA with batch annotation (pcRegscale)
   c. Correlation heatmap
6. Biological signal
   a. Correlation among protein complex members (complex_ks)
   b. Gene function prediction (func_auc)
   c. Sample class prediction (class_auc)
   d. PCA with sample class annotation
   e. Unsupervised clustering
7. Platform reproducibility (optional with QC sample)
   a. Coefficient of variation distribution (median CV)
8. Multi-omics concordance (optional with two omics)
   a. Gene-wise mRNA-protein correlation (gene wise cor)
   b. Sample-wise mRNA-protein correlation (sample wise cor)

# OmicsEV evaluation report

A few example evaluation reports are available at 
https://github.com/bzhanglab/OmicsEV.

# Session information

All software and respective versions used to produce this document are listed below.

```{r echo=FALSE}
sessionInfo()
```

# References


